---
title: "Drayage Simulation"
author: "Sean Fitzgerald"
date: "8/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```


```{r Load Libraries, include=FALSE}
library(geosphere)
library(tidyverse)
library(lubridate)
```

First, set up the initial parameters of the simulation. 
We randomly create a number of loads in one of the locations listed below, and add a drop off location in another one of these locations.
We randomly place trucks within the defined space of these locations.
```{r Setup Parameters for Simulation, echo=TRUE}

set.seed(1989)

# Iniital Parameters for simulation
numLoads <- 9
numTrucks <- 3

# All Loads originate at Port?
allStartPort = TRUE

# Pick up and Dropoff Locations of any and all loads
locations_df <- tibble(Locations = c("Port of Savannah", "Macon", "Jacksonville", "Savannah Warehouses and Rail", "Hilton Head"),
                       Latitude = c(32.119867, 32.810280, 30.353136, 32.081157, 32.213030),
                       Longitude = c(-81.139283, -83.564390, -81.627473, -81.145606, -80.729730))

# Date and Time of Load Windows
earliest_Pickup <- ymd_hms("2019/08/01 6:00:00")
latest_Pickup <- ymd_hms("2019/08/01 12:00:00")

earliest_Dropoff <- ymd_hms("2019/08/01 13:00:00")
latest_Dropoff <- ymd_hms("2019/08/01 20:00:00")

CarrierSpeed_MetersPerMinute <- 1609.34 # 60mph --> 1609 m/min.

CurrentTime_CarrierStart <- ymd_hms("2019/08/01 5:00:00")

```

Now we construct our "Load Board" - a listing of all the possible loads in the market.

```{r simulate Loads Data, echo=FALSE}

distanceMatrix <- locations_df %>% 
  select(-Locations)
distanceMatrix <- geosphere::distm(distanceMatrix[2:1], fun = distHaversine)
colnames(distanceMatrix) <- locations_df$Locations
rownames(distanceMatrix) <- locations_df$Locations
 


loads_df <- tibble(Load_ID = seq(1, numLoads),
                   Load_Weight = ceiling(rnorm(numLoads, mean = 100, sd = 15)))

if(allStartPort){
  # For all Trips originate at Port of Savannah
  trip_Sim_Matrix<- rbind(rep(x = 1, times = numLoads), 
                          replicate(numLoads,
                                    sample(x=seq(2, nrow(locations_df)), 1)))

} else {
  # For Random Geographic locations (start and end)
  trip_Sim_Matrix<- replicate(numLoads,
                              sample( x=seq(1, nrow(locations_df)),
                                      2,
                                      replace = F))
}

loads_df$Load_Origin_Location <- locations_df$Locations[trip_Sim_Matrix[1,]]
loads_df$Load_Destination_Location <- locations_df$Locations[trip_Sim_Matrix[2,]]
loads_df$LoadOrigin_Latitude <- locations_df$Latitude[trip_Sim_Matrix[1,]]
loads_df$LoadDestination_Latitude <- locations_df$Latitude[trip_Sim_Matrix[2,]]
loads_df$LoadOrigin_Longitude <- locations_df$Longitude[trip_Sim_Matrix[1,]]
loads_df$LoadDestination_Longitude <- locations_df$Longitude[trip_Sim_Matrix[2,]]

# Simulate pick up and dropoff windows
TimeWindowSim <- function(N, st=ymd_hms("2012/01/01 7:00:00"), et=ymd_hms("2012/01/01 9:00:00")) {
  require(lubridate)
     dt <- as.numeric(difftime(et,st,unit="mins"))
     ev <- ceiling(sort(runif(N,30, dt)))
     rt <- st + minutes(ev)
}

loads_df$PickupStartTime <- TimeWindowSim(numLoads, earliest_Pickup, latest_Pickup)


loads_df$DropoffStartTime <- TimeWindowSim(numLoads, earliest_Dropoff, latest_Dropoff)

loads_df <- loads_df %>% 
  mutate(PickupEndTime = TimeWindowSim(1, PickupStartTime, latest_Pickup),
         DropoffEndTime = TimeWindowSim(1, DropoffStartTime, latest_Dropoff))

# Calculate Trip distance. Will use straight line as proxy for trip time.
loads_df <- loads_df %>% 
  mutate(BundleDistance_meters = geosphere::distHaversine(p1 = cbind(LoadOrigin_Longitude, LoadOrigin_Latitude), p2 = cbind(LoadDestination_Longitude, LoadDestination_Latitude))) %>% 
  mutate(BundleDistance_meters = ifelse(allStartPort == T,
                                        2*BundleDistance_meters,
                                        BundleDistance_meters)) %>% 
  mutate(BundleTimeEstimate_minutes = ceiling(BundleDistance_meters/CarrierSpeed_MetersPerMinute)) #assumes 60mph converted to meters per minute

loads_df %>% 
  mutate(Load_ID = LETTERS[Load_ID]) %>% 
  knitr::kable(.)

```

Similarly, the below table shows the possible trucks vying for loads in the market.
```{r Simulate Truck Carriers data}


trucks_df <- tibble(Truck_ID = seq(1, numTrucks),
                    ELD_TimeRemaining = rpois(numTrucks, 7),
                    CurrentLocation_Longitude = runif(numTrucks, -83.564390, -81.139283),
                    CurrentLocation_Latitude = runif(numTrucks, 32.081157, 32.810280),
                    DesiredEndLocation_Longitude  = runif(numTrucks, -83.564390, -81.139283),
                    DesiredEndLocation_Latitude  = runif(numTrucks, 32.081157, 32.810280))
                    
trucks_df %>% 
  knitr::kable(.)
```

Better displayed on a map... 

```{r Leaflet of base starting conditions}

mapLocations_df <- tibble(ID = rep(x = c(loads_df$Load_ID, trucks_df$Truck_ID), 2),
                          lng = c(loads_df$LoadOrigin_Longitude, trucks_df$CurrentLocation_Longitude, loads_df$LoadDestination_Longitude, trucks_df$DesiredEndLocation_Longitude),
                          lat = c(loads_df$LoadOrigin_Latitude, trucks_df$CurrentLocation_Latitude, loads_df$LoadDestination_Latitude, trucks_df$DesiredEndLocation_Latitude),
                          description =  rep(c(paste0("Weight: ", loads_df$Load_Weight), paste0("ELD Hours Remaining: ", trucks_df$ELD_TimeRemaining)), 2),
                          position = c(rep("Start", times = nrow(loads_df)+nrow(trucks_df)), rep("End", times = nrow(loads_df)+nrow(trucks_df))),
                          group = rep(c(rep(x = "Load", times = nrow(loads_df)), rep(x = "Truck", times = nrow(trucks_df))), times = 2))

library(leaflet)

drayageIcons <- iconList(
  trucks = makeIcon("/home/sean/pCloudDrive/sfitzgerald Drive/Drayage/Map Icons/logistics-delivery-truck-in-movement.png", 32, 32),
  truckStop = makeIcon("/home/sean/pCloudDrive/sfitzgerald Drive/Drayage/Map Icons/icons8-sleeping-over-the-computer-64.png", 32, 32),
  loads = makeIcon("/home/sean/pCloudDrive/sfitzgerald Drive/Drayage/Map Icons/packages-transportation-on-a-truck.png", 32, 32),
  dropoff = makeIcon("/home/sean/pCloudDrive/sfitzgerald Drive/Drayage/Map Icons/Shipment_DropOff-512.png", 32, 32)
)

mapOfInterest <-  leaflet(mapLocations_df) %>% 
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(data = filter(mapLocations_df, group == "Load" & position == "Start"),
             label = ~LETTERS[ID],
             popup = ~description,
             clusterOptions = markerClusterOptions(),
             icon = ~drayageIcons$loads) %>% 
  addMarkers(data = filter(mapLocations_df, group == "Truck" & position == "Start"),
             label = ~ID,
             popup = ~description,
             clusterOptions = markerClusterOptions(),
             icon = ~drayageIcons$trucks) %>% 
  addMarkers(data = filter(mapLocations_df, group == "Load" & position == "End"),
             label = ~LETTERS[ID],
             popup = ~description,
             clusterOptions = markerClusterOptions(),
             icon = ~drayageIcons$dropoff) %>% 
  addMarkers(data = filter(mapLocations_df, group == "Truck" & position == "End"),
             label = ~ID,
             popup = ~description,
             clusterOptions = markerClusterOptions(),
             icon = ~drayageIcons$truckStop)

mapOfInterest
```


Check to see all the possible bundles (combinations of any two loads). A load ID of NA means no second load is taken (a bundle of 1 load). The table below only shows the first 20 rows as an example.

```{r check for potential bundles}

# add Load_ID_3 for 3-load bundle checks
bundles_df <- tibble(Load_ID = seq(0, numLoads),
                     Load_ID_2 = seq(0, numLoads)) %>% 
  complete(Load_ID, Load_ID_2) %>% 
  filter(Load_ID > 0) %>% 
  filter(Load_ID != Load_ID_2) %>%
  rowwise() %>% 
  # mutate(DeadHeadStartLocation = ifelse(Load_ID_2 > 0, 
  #                                       loads_df$Load_Destination_Location[Load_ID],
  #                                       "NA"),
  #        DeadHeadEndLocation = ifelse(Load_ID_2 > 0, 
  #                                       loads_df$Load_Origin_Location[Load_ID_2],
  #                                       "NA"))
  mutate(DeadHeadDistanceBtwnLoads_meters = ifelse(Load_ID_2 > 0,
                                          distHaversine(p1=cbind(loads_df$LoadDestination_Longitude[Load_ID],loads_df$LoadDestination_Latitude[Load_ID]),
                                                        p2=cbind(loads_df$LoadOrigin_Longitude[Load_ID_2], loads_df$LoadOrigin_Latitude[Load_ID_2])),
                                          0)) %>%
  mutate(BundleDistance_meters = ifelse(Load_ID_2 > 0,
                                        DeadHeadDistanceBtwnLoads_meters + loads_df$BundleDistance_meters[Load_ID] + loads_df$BundleDistance_meters[Load_ID_2],
                                        loads_df$BundleDistance_meters[Load_ID])) %>% 
  mutate(BundleTimeEstimate_minutes = ceiling(BundleDistance_meters/CarrierSpeed_MetersPerMinute)) 
  
viableBundles_df <- bundles_df %>% 
  rowwise() %>% 
  mutate(LastCall = ifelse(Load_ID_2 > 0,
                           loads_df$DropoffEndTime[Load_ID_2],
                           loads_df$DropoffEndTime[Load_ID])) %>% 
  filter(LastCall > loads_df$PickupStartTime[Load_ID] + minutes(BundleTimeEstimate_minutes))

viableBundles_df %>% 
  mutate(Load_ID = LETTERS[Load_ID]) %>% 
  mutate(Load_ID_2 = ifelse(Load_ID_2 > 0,
                            LETTERS[Load_ID_2],
                            NA)) %>% 
  head(., n = 20L) %>% 
  knitr::kable(.)
```

Pair each truck with each possible bundle (`r nrow(viableBundles_df)` possible bundles in this example) and check that the trip time will not exceed ELD requirements. The a small sample of the resulting trips are shown in the table below.

```{r Create Possble Trips}

possibleTrips_df <- expand.grid(Truck_ID = seq(1, numTrucks),
                        Load_ID = seq(1, numLoads),
                        Load_ID_2 = seq(0, numLoads)) 
   
keys <- c("Load_ID", "Load_ID_2")

trips_df <- merge(possibleTrips_df, viableBundles_df, by=keys) %>% 
  select(Truck_ID, Load_ID, Load_ID_2, everything()) %>% 
  rowwise() %>% 
  mutate(DistanceToFirstLoad_meters = distHaversine(p1 = cbind(trucks_df$CurrentLocation_Longitude[Truck_ID], trucks_df$CurrentLocation_Latitude[Truck_ID]), 
                                         p2 = cbind(loads_df$LoadOrigin_Longitude[Load_ID], loads_df$LoadOrigin_Latitude[Load_ID]))) %>% 
  mutate(TimeToFirstLoad_minutes = ceiling(DistanceToFirstLoad_meters/CarrierSpeed_MetersPerMinute)) %>% 
  filter(CurrentTime_CarrierStart + minutes(TimeToFirstLoad_minutes) < loads_df$PickupEndTime[Load_ID]) %>% 
  mutate(EstimatedEndOfTripTime = CurrentTime_CarrierStart + minutes(TimeToFirstLoad_minutes) + BundleTimeEstimate_minutes) %>% 
  filter(ifelse(Load_ID_2 > 0,
                EstimatedEndOfTripTime < loads_df$DropoffEndTime[Load_ID_2],
                EstimatedEndOfTripTime < loads_df$DropoffEndTime[Load_ID])) %>% 
  mutate(TripHome_meters = ifelse(Load_ID_2 > 0, 
                                  distHaversine(p1 = cbind(loads_df$LoadDestination_Longitude[Load_ID_2],
                                                                          loads_df$LoadDestination_Latitude[Load_ID_2]),
                                                               p2 = cbind(trucks_df$DesiredEndLocation_Longitude[Truck_ID],
                                                                          trucks_df$DesiredEndLocation_Latitude[Truck_ID])),
                                  distHaversine(p1 = cbind(loads_df$LoadDestination_Longitude[Load_ID],
                                                                          loads_df$LoadDestination_Latitude[Load_ID]),
                                                               p2 = cbind(trucks_df$DesiredEndLocation_Longitude[Truck_ID],
                                                                          trucks_df$DesiredEndLocation_Latitude[Truck_ID])))) %>% 
  mutate(TripHome_minutes = ceiling(TripHome_meters / CarrierSpeed_MetersPerMinute)) %>% 
  filter(EstimatedEndOfTripTime + minutes(TripHome_minutes) < CurrentTime_CarrierStart + hours(trucks_df$ELD_TimeRemaining[Truck_ID])) %>% 
  mutate(TotalTripDistance = TripHome_meters + DistanceToFirstLoad_meters + BundleDistance_meters,
         TotalTripTime = ceiling(TripHome_minutes + TimeToFirstLoad_minutes + BundleTimeEstimate_minutes))

trips_df %>% 
  mutate(Load_ID = LETTERS[Load_ID]) %>% 
  mutate(Load_ID_2 = ifelse(Load_ID_2 > 0,
                            LETTERS[Load_ID_2],
                            NA)) %>% 
  head(., n = 10L) %>% 
  knitr::kable(.)
```

Now, Optimize the trip each truck should take. Using a genetic algorithm, the optimization routine finds the routes for each truck that carry the most loads while minimizing overall trip time, including trucks going to a desired end location after dropping off their last load.
```{r Optimize, include=FALSE}

# Optimize based on Time

fitFunction <- function(x) {
  x <- as.logical(x)
  
  potentialLoads <- unique(c(trips_df$Load_ID, trips_df$Load_ID_2))
  # filter out the 0 loads
  potentialLoads <- potentialLoads[potentialLoads>0]
  
  loadsCarried <- c(trips_df$Load_ID[x], trips_df$Load_ID_2[x])
  # filter out the 0 loads
  loadsCarried <- loadsCarried[loadsCarried>0]
  
  
  # Define fitness as positive result of carrying a load
  fitness = sum(potentialLoads %in% loadsCarried)*max(trips_df$TotalTripTime)
  
  ## Penalties
  # Total Trip time (we are trying to minimize this)
  penalty_Time <- -1*sum(trips_df$TotalTripTime[x])
  
  # Penalty for leaving loads behind (otherwise solution is to take 1 load)
  penalty_NotCarryingLoads <- -1*sum(length(potentialLoads) - unique(loadsCarried))*max(trips_df$TotalTripTime)     
  
  # Harsh penalties for breaking constraints
  penalty_ReuseTrucks  <- ifelse(any(duplicated(trips_df$Truck_ID[x])),
                                 -1000*sum(duplicated(trips_df$Truck_ID[x]))*max(trips_df$TotalTripTime),
                                 0)
  
  penalty_ReuseLoads  <- ifelse(any(duplicated( c(trips_df$Load_ID[x], trips_df$Load_ID_2[x]) )),
                                 -1000*sum(duplicated( c(trips_df$Load_ID[x], trips_df$Load_ID_2[x]) ))*max(trips_df$TotalTripTime),
                                 0)
  
  fitness = fitness + penalty_Time + penalty_ReuseTrucks + penalty_ReuseLoads
  
  return(fitness)
}


library(GA)

ga_results <- ga(type = "binary",
                 fitness = fitFunction,
                 lower = rep(x = 0, nrow(trips_df)),
                 upper = rep(x = 1, nrow(trips_df)),
                 nBits = nrow(trips_df),
                 maxiter=1000, # Maximum number of generations
                 run=100,     # Stop if the best-so-far fit hasn't improved for 'run' generations 
                 popSize=100, 
                 seed=101)
                 
```
The results of the optimization are plotted below. The curve should converge to a flat horizontal line at the top. This means the optimizer is confident and consistent in the answer:

```{r PLot GA results, echo=FALSE}
  plot(ga_results)
```


The resulting table of optimized trips:
```{r Show Optimized Trips}
solution <- (summary(ga_results)$solution)

trips_df$isOptimal <- solution[1:nrow(trips_df)]

optimizedTrips <- trips_df %>% filter(isOptimal == 1)

optimizedTrips %>% 
  mutate(Load_ID = LETTERS[Load_ID]) %>% 
  mutate(Load_ID_2 = ifelse(Load_ID_2 > 0,
                            LETTERS[Load_ID_2],
                            NA)) %>% 
  knitr::kable(.)
```


Let's show the routes mapped out.  Each color line is a different truck's path from their current position to their desired end location.

```{r Add optimized trip lines}

# positionVector = c("CurrentLocation", "LoadOrigin", "LoadDestination", "LoadOrigin", "LoadDestination", "DesiredEndLocation")

optimizedTrips_extended_df <- optimizedTrips %>% 
  select(Truck_ID, Load_ID, Load_ID_2) %>% 
  gather(key = "LoadOrder", value = "Load_ID", -Truck_ID) %>% 
  mutate(LoadOrder = ifelse(str_detect(string = LoadOrder, pattern = "_2"), 2, 1)) %>% 
  left_join(., trucks_df, by= "Truck_ID") %>% 
  left_join(., loads_df, by="Load_ID") %>% 
  select(Truck_ID, Load_ID, LoadOrder, CurrentLocation_Longitude, CurrentLocation_Latitude, LoadOrigin_Longitude, LoadOrigin_Latitude, LoadDestination_Longitude,
         LoadDestination_Latitude, DesiredEndLocation_Longitude, DesiredEndLocation_Latitude) %>% 
  gather(key = TripNode, value = Coordinate, -c(ends_with("ID"), LoadOrder)) %>% 
  separate(col = TripNode, sep = "_", into = c("TripNode", "CoordinateAxis")) %>% 
  spread(key = CoordinateAxis, value = Coordinate) %>% 
  mutate(LoadOrder = ifelse(TripNode == "LoadOrigin" & LoadOrder == 2,
                            3,
                            ifelse(TripNode == "LoadDestination",
                            2*LoadOrder,
                            LoadOrder))) %>% 
  mutate(LoadOrder = ifelse(TripNode == "CurrentLocation", 
                            0,
                            ifelse(TripNode == "DesiredEndLocation",
                                   5,
                                   LoadOrder))) %>%  
  filter(!duplicated(.)) %>% 
  group_by(Truck_ID) %>% 
  filter(duplicated(TripNode == "CurrentLocation")) %>% 
  # filter(duplicated(TripNode == "DesiredEndLocation")) %>%
  arrange(Truck_ID, LoadOrder) %>% 
  ungroup() %>% 
  select(Truck_ID, Latitude, Longitude)
  
  
# Map the routes

pal <- colorFactor(
  palette = "viridis",
  domain = optimizedTrips$Truck_ID
)
routeMap <- mapOfInterest 
for (i in unique(optimizedTrips_extended_df$Truck_ID)) {
  routeMap <- routeMap %>% 
    addPolylines(data = subset(optimizedTrips_extended_df, Truck_ID == i),
                 group = ~Truck_ID,
                 lng = ~Longitude,
                 lat = ~Latitude,
                 color = ~pal(Truck_ID[i]))
  
}

routeMap <- routeMap %>% 
  addLegend(position = "bottomleft",
            pal = pal,
            values = ~optimizedTrips_extended_df$Truck_ID,
            title = "Truck ID Route",
            opacity = 1)

routeMap
```


Finally, the below table shows the remaining loads that were not able to be carried by the trucks. We can run this process again on just these loads with a new set of trucks (say RoadOne's own vehicles) to find new bundles and optimize their trips.
```{r}

remainingLoads_df <- loads_df %>% 
  filter(!(Load_ID %in% optimizedTrips$Load_ID)) %>% 
  filter(!(Load_ID %in% optimizedTrips$Load_ID_2)) %>% 
  mutate(Load_ID = LETTERS[Load_ID])


remainingLoads_df %>% 
    knitr::kable(.)

```

