---
title: "Single Lane Optimization"
output: html_notebook
---

# Day-ahead Auction Resolution/Optimization

This notebook solves test auction scenarios. 
TO DO List: 
1. Implement Load-specific ceiling prices. -- Merge all data into 1 df
2. Standardize Input and output format and make accessible
3. Check assumptions about the number of Bids == Loads == Carry Capacity--> NO bids are based on capcity of the carrier and not the number of loads in the lane
4. add simulation of data
5. stress test and test different scenarios
 

```{r load Libraries, message=FALSE, warning=FALSE, include=FALSE}

library(tidyverse)
library(purrr)

set.seed(1988)

```


```{r Simulate Loads}

# Parameters for sampling

numLanes = 1
numLoadsPerLane = c(10, 10, 10)
#  OR
MaxNumLoadsPerLane = 10
numLoadsPerLane = sample(x = seq(1, 10), size = numLanes, replace = T)
```

## Load the example data of 3 Lanes and 3 Carriers

Load Bids by lane, carrier, and load
```{r Load Bids example, echo=FALSE}

# Bids_df <- read_csv("/home/flai/pCloudDrive/Data/DadsDrayExample.csv") %>% 
Bids_df <- read_csv("/home/sean/pCloudDrive/Data/ShippingContainerExample.csv") %>% 
  gather(key = Loads, value = Bid_Price, -c(Carrier, Lanes)) %>%
  # mutate(Bid_Price = ifelse(Bid_Price == 0, 9999*max(Bid_Price), Bid_Price)) %>% 
  filter(Bid_Price != 0) %>% 
  rename(Lane = Lanes,
         numLoadsBid = Loads) %>% 
  mutate(numLoadsBid = as.numeric(numLoadsBid)) %>% 
  group_by(Lane) %>% 
  mutate(BidID = 1:n()) %>% 
  ungroup()

Bids_df

Bids_df_byLane <- Bids_df %>% group_by(Lane) %>% group_split()
```

Simulate the Loads and ceiling prices for each lane
```{r Load or Simulate Lane Data, echo=FALSE}
  Lanes_df <- expand.grid(Lane = seq(1,numLanes),
                          LoadID = seq(1, MaxNumLoadsPerLane)) %>% 
  mutate(., Ceiling_Price = sample(x = c(540, 390, 550, 500, 525, 475), size = nrow(.), replace = T)) %>% 
  group_by(Lane) %>% 
  ungroup()

# Lanes_df$Ceiling_Price <- 500
# Lanes_df %>% group_by(Lane) %>% summarise(mean(Ceiling_Price)) #when Individual loads in a lane have different ceiling prices
Loads_df_byLane <- Lanes_df %>% group_by(Lane) %>% group_split()
Loads_df_byLane
```



```{r Merge Data Sets}

Merged_df <- Bids_df %>% 
  left_join(., Lanes_df) 

```


## Objective Function
Decide what strategy or objective we are trying to optimize. Optimize with ceiling constraint or maximize number of loads taken and set the ceiling price (currently the same ceiling price across all lanes)

```{r Pre-opt parameters}
# maxNumLoads_setting <- F
# If param is set to T, maximize number of loads by checking ceiling price average per load across the net for the entire lane
# If param is set to F, each bid considered must have a bid price per load below the ceiling
```

```{r Preprocessing}

numberLanes <- n_distinct(Lanes_df$Lane)
numCarriers <- n_distinct(Bids_df$Carrier)
numBids_byLane <- Merged_df %>% group_by(Lane) %>% group_size()

MergedData_byLane <- Merged_df %>% mutate(CP_Delta = Ceiling_Price - Bid_Price) %>% group_by(Lane) %>% group_split()

```

## Approach
Treat every lane as a separate auction! Each number of loads from a carrier is a separate bid, with only 1 bid acceptable per carrier. 

### lpSolve Algorithm

```{r Individual Load Pricing Model}
library(lpSolveAPI)

  weight_Profit <- .01
  
  for (lane in seq(1, numberLanes)) {
  
  numBid_LoadCombinations <-  numBids_byLane[[lane]]
  numLoads <- nrow(Loads_df_byLane[[lane]])
  numBids <- max(MergedData_byLane[[lane]]$BidID)
  lprec <- NULL
  lprec <- make.lp(0, numBid_LoadCombinations+numBids)
  # Set the objective function
  set.objfn(lprec, 
            obj = c(weight_Profit*MergedData_byLane[[lane]]$CP_Delta,
                    Bids_df_byLane[[lane]]$numLoadsBid))
  
  # add constraints
  # loads
  
  # Each Load can only be used 1 time
  for (load in seq(1, numLoads)) {
    add.constraint(lprec,
                   xt = rep(x = 1,
                            times = sum(MergedData_byLane[[lane]]$LoadID == load)),
                   indices = which(MergedData_byLane[[lane]]$LoadID == load),
                   "<=", rhs = 1)
  }
  
  # Global sum of loads used must be <= total number of loads available in the lane
  add.constraint(lprec,
                 xt = rep(x = 1,
                          times = numBid_LoadCombinations),
                 indices = 1:numBid_LoadCombinations,
                 "<=", rhs = nrow(Loads_df_byLane[[lane]]))
  
  add.constraint(lprec,
                 xt = Bids_df_byLane[[lane]]$numLoadsBid,
                 indices = (numBid_LoadCombinations+1):(numBid_LoadCombinations+numBids),
                 "<=", rhs = nrow(Loads_df_byLane[[lane]]))
  
  # carriers
  # only 1 bid per carrier
  for (carrier in seq(1, numCarriers)) {
    add.constraint(lprec,
            xt = rep(x = 1,
                     times = sum(Bids_df_byLane[[lane]]$Carrier == unique(Bids_df_byLane[[lane]]$Carrier)[[carrier]])),
            indices = numBid_LoadCombinations+ which(Bids_df_byLane[[lane]]$Carrier == unique(Bids_df_byLane[[lane]]$Carrier)[[carrier]]),
            "<=", rhs = 1)
  }
  
  # bid amounts == loads carried
  for (bid in seq(1, numBids)) {
    add.constraint(lprec,
            xt = 1,
            indices = numBid_LoadCombinations+bid,
            "<=", rhs = 1)
    
    add.constraint(lprec,
            xt = c(rep(x=(1/Bids_df_byLane[[lane]]$numLoadsBid[[bid]]), times = sum(MergedData_byLane[[lane]]$BidID == bid)),
                   -1),
            indices = c(which(MergedData_byLane[[lane]]$BidID == bid),numBid_LoadCombinations+bid),
            "=", rhs = 0)
    
    add.constraint(lprec,
            xt = c(rep(x=1, times = sum(MergedData_byLane[[lane]]$BidID == bid)),
                   -Bids_df_byLane[[lane]]$numLoadsBid[[bid]]),
            indices = c(which(MergedData_byLane[[lane]]$BidID == bid),numBid_LoadCombinations+bid),
            "=", rhs = 0)
  }
  
  
  #Make decision variables binary and integer semi-continuous
  # set.type(lprec, 1:numBid_LoadCombinations, type = "binary")
  set.type(lprec, columns = 1:(numBid_LoadCombinations+numBids), type = "binary") # set all decision vars to binary

  # set.type(lprec, (numBid_LoadCombinations+1):(numBid_LoadCombinations+numBids), type = "integer")
  # set.semicont(lprec, (numBid_LoadCombinations+1):(numBid_LoadCombinations+numBids))
  # set.bounds(lprec,
  #            lower = Bids_df_byLane[[lane]]$numLoadsBid,
  #            upper = Bids_df_byLane[[lane]]$numLoadsBid,
  #            columns = (numBid_LoadCombinations+1):(numBid_LoadCombinations+numBids))

  # Set to maximize or minimize depending on maxNumLoads_setting
  lp.control(lprec, sense='max') 
  
  # Solve the model
  solve(lprec)
  
  # Record Results
  MergedData_byLane[[lane]]$UseBid <- get.variables(lprec)[1:numBid_LoadCombinations]
  # Lanes_df$TotalSavings[lane] <-
    get.objective(lprec)
  # Lanes_df$LoadsMoved[lane] <- 
    # sum(MergedData_byLane[[lane]]$UseBid*MergedData_byLane[[lane]]$Loads)
 
  }
```


Results of the optimization have been written to "r savename". Sample Table of the Bid Accept/Reject decisions is below

```{r Auction Results, echo=FALSE}

Results
```


Aggregate summary of the auction outcomes by lane are shown in the following table.

```{r Show Lane Summaries, echo=FALSE}
Lanes_df
```




### Evolutionary/Genetic Algorithm

This algorithm will be more useful for the "fuzzy" variable of forecasted supply/demand or for maximizing both quantity and savings.


```{r GA, eval=FALSE, include=FALSE}

library(GA) 

fitness <- function(x) {
  
  Profit <- weight_Profit*MergedData_byLane[[lane]]$CP_Delta
  
  NumLoadsCarried = sum(MergedData_byLane[[lane]]$numLoadsBid)
  
  
  fitness <- w_profit*Profit + w_loads*NumLoadsCarried
  
  return(fitness)
}

```


